### 2019-12-05 paper6/paper.tex

%------------------------------------------------------------------------------
% Define title, author(s), affiliation and publishing status
%
\papertitle[Effective slip over partially filled microcavities] % Short title used in headlines (optional)
{%
  Effective slip over partially filled microcavities and its possible failure% THE COMMENT SYMBOL AT THE END OF THIS LINE IS NEEDED
}%
%
\papertoctitle{Effective slip over partially filled microcavities and its possible failure} % Title for toc
%
\paperauthor[Ge, Holmgren, Kronbichler, Brandt, Kreiss] % Short authors used in headlines and List Of Papers
{%
  Zhouyang Ge$^1$, Hanna Holmgren$^2$, Martin Kronbichler$^3$, \\Luca Brandt$^1,$, Gunilla Kreiss$^2$%
}%
%
\listpaperauthor{Z. Ge, H. Holmgren, M. Kronbichler, L. Brandt, G. Kreiss}% (optional) Short authors used in List Of Papers
%
\paperaffiliation
{%
  $^1$ Linn\'e FLOW Centre and SeRC, KTH Mechanics, S-100 44 Stockholm, Sweden\\%
  $^2$ Division of Scientific Computing, Department of Information Technology, \\Uppsala University, Box 337, 751 05 Uppsala, Sweden\\%
  $^3$ Institute for Computational Mechanics, Technical University of Munich, Boltzmannstrasse 15, 85748 Garching b. München, Germany%
}%
%
\paperjournal[Phys. Rev. Fluids] % Short publish info used in List Of Papers
{%
	Physical Review Fluids%
}%
%
\papervolume{3}%
%
\papernumber{054201}%
%
\paperpages{}%
%
\paperyear{2018}%
%
\papersummary%
{% Insert summary of the paper here (used in introduction)
\todo
}%
%
\graphicspath{{paper6/}}%
%
%
%===============================================================================
%                            BEGIN PAPER
%===============================================================================
%
\begin{paper}

\makepapertitle

%------------------------------------------------------------------------------
% Abstract
%------------------------------------------------------------------------------
%
\begin{paperabstract}
\Ge{Motivated by the emerging applications of liquid-infused surfaces (LIS), we study the drag reduction and robustness of transverse flows over two-dimensional microcavities partially filled with an oily lubricant.
Using separate simulations at different scales, characteristic contact line velocities at the fluid-solid intersection are first extracted from nano-scale phase field simulations and then applied to micron-scale two-phase flows, thus introducing a multiscale numerical framework to model the interface displacement and deformation within the cavities. 
As we explore the various effects of the lubricant-to-outer-fluid viscosity ratio $\tilde{\mu}_2/\tilde{\mu}_1$, the capillary number Ca, the static contact angle $\theta_s$, and the filling fraction of the cavity $\delta$, we find that the effective slip is most sensitive to the parameter  $\delta$. The effects of $\tilde{\mu}_2/\tilde{\mu}_1$ and $\theta_s$ are generally intertwined, but weakened if $\delta < 1$.
Moreover, for an initial filling fraction $\delta =0.94$, our results show that the effective slip is nearly independent of the capillary number, when it is small. Further increasing Ca to about $0.01 \tilde{\mu}_1/\tilde{\mu}_2$, we identify a possible failure mode, associated with lubricants draining from the LIS, for $\tilde{\mu}_2/\tilde{\mu}_1 \lesssim 0.1$. Very viscous lubricants (\eg $\tilde{\mu}_2/\tilde{\mu}_1 >1$), on the other hand, are immune to such failure due to their generally larger contact line velocity.}
\end{paperabstract}

%------------------------------------------------------------------------------
% Article
%------------------------------------------------------------------------------
%
\input{paper6/article.tex}

%------------------------------------------------------------------------------
% Bibliography
%------------------------------------------------------------------------------
%
\clearpage
\bibliographystyle{jfm}
\bibliography{thesis}
%
\IfFileExists{paper6/paper.bbl}{\input{paper6/paper.bbl}}{}

%===============================================================================
%                            END PAPER
%===============================================================================
\end{paper}