### 2019-12-14 frontmatter.tex
conference/workshop info added

%-------------------------------------------------------------------------------
% Cover page
%-------------------------------------------------------------------------------
%
\makecoverpage

%-------------------------------------------------------------------------------
% Dedication (comment if none)
%-------------------------------------------------------------------------------
%
\dedication{
	{\it If the scientist had an infinity of time at his disposal, it would be sufficient to say to him, ``Look, and look carefully.''}\\[20pt]
	Henri Poincar\'e
}

%-------------------------------------------------------------------------------
% Abstract (English)
%-------------------------------------------------------------------------------
%
% COMMENT: the abstract should be limited to 2000 characters and the keywords
% to 250 characters because otherwise they will not fit in the mailing sheet.
%
\begin{abstract}
	Access points  must work. In fact, few cryptographers would disagree with
	the emulation of robots, which embodies the technical principles of
	complexity theory. Hyp, our new application for lossless information, is the
	solution to all of these challenges.
		%
		\keywords{super-lasers, space-stations, The Force, cookies.}
		%
\end{abstract}

%-------------------------------------------------------------------------------
% Abstrakt (Swedish) 
%-------------------------------------------------------------------------------
%
% COMMENT: the abstract should be limited to 2000 characters and the keywords
% to 250 characters because otherwise they will not fit in the mailing sheet.
%
\begin{abstrakt}
	\r{A}tkomstpunkter m\r{a}ste arbeta. I sj\"{a}lva verket skulle f\r{a}
	kryptografer h\r{a}ller inte med emulering av robotar, som
	f\"{o}rkroppsligar de tekniska principer av komplexitetsteori. Hyp, v\r{a}r
	nya applikation f\"{o}r f\"{o}rlustfri information är l\"{o}sningen på alla
	dessa utmaningar.
		%
		\nyckelord{super-lasrar, rymdstationer, Kraften, kakor.}
		%
\end{abstrakt}

%-------------------------------------------------------------------------------
% Preface page
%-------------------------------------------------------------------------------
%
\begin{preface}
	This thesis summarizes a selection of studies on droplet interactions and suspension flows
        that may find applications in microfluidics, rheology, and soft matter physics.
        A brief introduction of some basic physical concepts and numerical methods is presented in Part I,
        followed by eight journal articles or preprints in Part II, see below.
        For consistency, all papers are formatted in the \texttt{jfm} style as the rest of the thesis.
        Their contents remain faithful to the original publications.
\end{preface}

%-------------------------------------------------------------------------------
% Division of work between authors
%-------------------------------------------------------------------------------
%
\begin{divisionofwork}
	The main thesis advisor is Professor Luca Brandt (LB).
	Associate Professor Outi Tammisola (OT) acts as the co-advisor.

	\paperitem
                LB directed the research.
                Z. Ge (ZG) and J-Ch. Loiseau developed, implemented, and validated the methods.
                OT oversaw the entire project.
		ZG wrote the paper with inputs from all other authors.

	\paperitem
                A. Leshansky (AL) directed the research.
                I. Fouxon (IF) derived the theory.
		ZG performed the simulation.
                IF wrote the paper with inputs from the rest of the authors.

	\paperitem
                Following Paper 1, ZG continued the research, performed the simulation,
                and wrote the paper with inputs from the rest of the authors.

	\paperitem
                AL directed the research.
                IF constructed the theory.
		B. Rubinstein performed the multipole expansion simulation.
                ZG performed the Navier-Stokes simulation.
                IF wrote the paper with inputs from all other authors.

	\paperitem
                LB conceived the project.
                M.E. Rosti (MER) implemented the method and performed the simulation.
		ZG performed the droplet statistics and scaling analyses.
                S.S. Jain performed the droplet rendering.
                M.S. Dodd performed the turbulent kinetic energy analysis.
                MER wrote the paper with inputs from the rest of the authors.

	\paperitem
                G. Kreiss initiated the project.
                H. Holmgren implemented the code and performed the preliminary Stokes simulation.
                M. Kronbichler performed the phase field simulation.
		ZG performed the final Stokes simulation, and 
                wrote the paper with inputs from the rest of the authors.

	\paperitem
                M. Minale and LB directed the research.
                R. Martone and C. Carotenuto performed the experiment.
                R. Radhakrishnan provided part of the simulation data.
		ZG performed the main simulation and wrote the paper with inputs from the rest of the authors.

	\paperitem
                LB led the research.
		ZG developed and implemented the method, with inputs from OT.
                ZG wrote the paper.

\end{divisionofwork}

%-------------------------------------------------------------------------------
% Additional publications (comment if none)
%-------------------------------------------------------------------------------
%
%\begin{otherpublications}
%	The following papers, although related, are not included in this thesis.
%
%  \paperitem%
%    {Anakin Skywalker \& Master Obi-Wan Kenobi}% Authors
%    {3639}% Year
%    {The light sabre: an elegant weapon for a more civilized age}% Title
%    {Jedi Journal of Weapons}% Journal
%    {33}% Volume
%    {2}% Number
%    {pp. 55--60}% Pages
%
%\end{otherpublications}

%-------------------------------------------------------------------------------
% Conferences (comment if none)
%-------------------------------------------------------------------------------
%
%\clearpage
\begin{conferences}
        In addition to journal publications, part of the work in this thesis were
        presented in the following conferences and workshops. The presenting author is underlined.

  \conferenceitem%
    {\underline{Z. Ge}, L. Brandt}% Authors
    {2016}% Year
    {Simulation of the self-assembly of colloidal particles in a microchannel}% Title
    {7th Summer school of Complex Motions in Fluids}% Conference
    {Zenderen, Twente, Netherlands}% Location

  \conferenceitem%        
    {Z. Ge, \underline{O. Tammisola}, J.Ch. Loiseau, L. Brandt}% Authors
    {2016}% Year
    {Direct numerical simulation of the self-assembly of colloidal particles in a micro-channel}% Title
    {1$^{st}$ International Conference on Multiscale Applications of Surface Tension}% Conference
    {Brussel, Belgium}% Location

  \conferenceitem%
    {\underline{Z. Ge}, L. Brandt}% Authors
    {2016}% Year
    {Do self-assembly colloidal droplets behave like droplets?}% Title
    {69$^{th}$ Annual Meeting of the APS Division of Fluid Dynamics}% Conference
    {Portland, Oregon, USA}% Location

  \conferenceitem%
    {\underline{Z. Ge}, O. Tammisola, L. Brandt}% Authors
    {2017}% Year
    {A fast mass-preserving interface-correction level set/ghost fluid method for colloidal suspensions under depletion forces}% Title
    {3$^{rd}$ International Conference on Numerical Methods of Multiphase Flows (ICNMMF-III)}% Conference
    {Tokyo, Japan}% Location

  \conferenceitem%
    {\underline{Z. Ge}, H. Holmgren, M. Kronbichler, L. Brandt, G. Kreiss}% Authors
    {2018}% Year
    {Effective slip over partially filled microcavities and its possible failure}% Title
    {12$^{th}$ European Fluid Mechanics Conference}% Conference
    {Vienna, Austria}% Location

  \conferenceitem%
    {\underline{Z. Ge}, O. Tammisola, L. Brandt}% Authors
    {2018}% Year
    {Flow-assisted droplet assembly in a 3D microfluidic channel}% Title
    {71$^{st}$ Annual Meeting of the APS Division of Fluid Dynamics}% Conference
    {Atlanta, Georgia, USA}% Location

  \conferenceitem%
    {\underline{Z. Ge}, O. Tammisola, L. Brandt}% Authors
    {2018}% Year
    {Flow-assisted droplet assembly in a 3D microfluidic channel}% Title
    {Scattering and Dynamics of Flowing Soft Material}% Conference
    {Lund, Sweden}% Location

  \conferenceitem%
    {\underline{Z. Ge}, O. Tammisola, L. Brandt}% Authors
    {2019}% Year
    {Flow-assisted droplet assembly in a 3D microfluidic channel}% Title
    {Colloidal Science \& Metamaterials (CSM)}% Conference
    {Paris, France}% Location

\end{conferences}

%-------------------------------------------------------------------------------
% Table of contents
%-------------------------------------------------------------------------------
%
\tableofcontents