### 2020-01-26 paper1/paper.tex

%------------------------------------------------------------------------------
% Define title, author(s), affiliation and publishing status
%
\papertitle[ICLS/GFM for droplet suspensions under depletion forces] % Short title used in headlines (optional)
{%
  An efficient mass-preserving interface-correction level set/ghost fluid method
  for droplet suspensions under depletion forces% THE COMMENT SYMBOL AT THE END OF THIS LINE IS NEEDED
}%
%
\papertoctitle{An efficient mass-preserving interface-correction level set/ghost fluid method
  for droplet suspensions under depletion forces} % Title for toc
%
\paperauthor[Ge, Loiseau, Tammisola, Brandt] % Short authors used in headlines and List Of Papers
{%
  Zhouyang Ge, Jean-Christophe Loiseau, Outi Tammisola, Luca Brandt%
}%
%
\listpaperauthor{Z. Ge, J-Ch. Loiseau, O. Tammisola, L. Brandt}% (optional) Short authors used in List Of Papers
%
\paperaffiliation
{%
  Linn\'e FLOW Centre and SeRC, KTH Mechanics, S-100 44 Stockholm, Sweden%
}%
%
\paperjournal[J. Comput. Phys.] % Short publish info used in List Of Papers
{%
	Journal of Computational Physics%
}%
%
\papervolume{353}%
%
\papernumber{}%
%
\paperpages{435--459}%
%
\paperyear{2018}%
%
\papersummary%
{% Insert summary of the paper here (used in introduction)
In this first work, we develop and implement an interface-correction level set/ghost fluid method (\texttt{ICLS/GFM})
for high-fidelity simulations of liquid droplets in an immiscible carrier fluid.
The main novelty is a global mass correction scheme that is conceptually simple,
relatively straightforward to program,
and efficient since the scheme does not need to be excuted at every time step.
The method works best for well-resolved droplets as validated in the paper;
however, extra care must be taken when applying it otherwise.
In the same paper, we also formulate a hydrodynamic model for surfactant-induced depletion forces between neighbouring drops,
removing the need to simulate suspending surfactants,
hence further improving the computational efficiency.
The overall methodology can be readily used to study droplet interactions in liquid flows,
and may be extended to simulate multiphase flows in porous media in general.
}%
%
\graphicspath{{paper1/}}%
%
%
%===============================================================================
%                            BEGIN PAPER
%===============================================================================
%
\begin{paper}

\makepapertitle

%------------------------------------------------------------------------------
% Abstract
%------------------------------------------------------------------------------
%
\begin{paperabstract}
Aiming for the simulation of colloidal droplets in microfluidic devices, we present here a numerical method for two-fluid systems subject to surface tension and depletion forces among the suspended droplets. The algorithm is based on an efficient solver for the incompressible two-phase Navier-Stokes equations, and uses a mass-conserving level set method to capture the fluid interface. 
The four novel ingredients proposed here are, firstly, an interface-correction level set (ICLS) method; % that efficiently preserves mass. 
global mass conservation is achieved by performing an additional advection near the interface, with a correction velocity obtained by locally solving an algebraic equation, which is easy to implement in both 2D and 3D. 
Secondly, we report a second-order accurate geometric estimation of the curvature at the interface and, 
thirdly, the combination of the ghost fluid method with the fast pressure-correction approach enabling an accurate and fast computation even for large density contrasts. 
Finally, we derive %sharp formulation for the pressure difference 
a hydrodynamic model for the interaction forces
induced by depletion of surfactant micelles and combine it with a multiple level set approach to study short-range interactions among droplets in the presence of attracting forces.
        \keywords{Multiphase flow, Level set method, Ghost fluid method, Droplet, Depletion force}
\end{paperabstract}

%------------------------------------------------------------------------------
% Article
%------------------------------------------------------------------------------
%
\input{paper1/article.tex}

%------------------------------------------------------------------------------
% Bibliography
%------------------------------------------------------------------------------
%
%\clearpage
\bibliographystyle{jfm}
\bibliography{thesis}
%
\IfFileExists{paper1/paper.bbl}{\input{paper1/paper.bbl}}{}

%===============================================================================
%                            END PAPER
%===============================================================================
\end{paper}