### 2020-01-26 paper2/paper.tex

%------------------------------------------------------------------------------
% Define title, author(s), affiliation and publishing status
%
\papertitle[Flow-assisted droplet assembly in a 3D microfluidic channel] % Short title used in headlines (optional)
{%
  Flow-assisted droplet assembly in a 3D microfluidic channel% THE COMMENT SYMBOL AT THE END OF THIS LINE IS NEEDED
}%
%
\papertoctitle{Flow-assisted droplet assembly in a 3D microfluidic channel} % Title for toc
%
\paperauthor[Ge, Tammisola, Brandt] % Short authors used in headlines and List Of Papers
{%
  Zhouyang Ge, Outi Tammisola, Luca Brandt%
}%
%
\listpaperauthor{Z. Ge, O. Tammisola, L. Brandt}% (optional) Short authors used in List Of Papers
%
\paperaffiliation
{%
  Linn\'e FLOW Centre and SeRC, KTH Mechanics, S-100 44 Stockholm, Sweden%
}%
%
\paperjournal[Soft Matter.] % Short publish info used in List Of Papers
{%
	Soft Matter%
}%
%
\papervolume{15}%
%
\papernumber{}%
%
\paperpages{3451--3460}%
%
\paperyear{2019}%
%
\papersummary%
{% Insert summary of the paper here (used in introduction)
In this paper, we perform a series of simulations, using methods developed in \emph{Paper 1},
to examine the droplet interaction in increasingly realistic conditions, mimicking a microfluidic experiment.
The main objective is to elucidate the physical mechanism that leads to the experimentally observed droplet clustering,
which had been explained by a dipolar interaction theory.
Contrary to previous belief, we find that dipolar interactions alone cannot account for the fast droplet dynamics;
a number of 3D effects must be included to faithfully reproduce the observation.
This conclusion is supported by theoretical derivations of particle interactions in the Stokes flow (\emph{Papers 3 \& 4}),
implying that designing optimal microfluidic channels to directly produce large droplet crystals remains difficult in practice.
}%
%
\graphicspath{{paper2/}}%
%
%
%===============================================================================
%                            BEGIN PAPER
%===============================================================================
%
\begin{paper}

\makepapertitle

%------------------------------------------------------------------------------
% Abstract
%------------------------------------------------------------------------------
%
\begin{paperabstract}
Self-assembly of soft matter, such as droplets or colloids, has become a promising scheme to engineer novel materials,
model living matter, and explore non-equilibrium statistical mechanics.
In this article, we present detailed numerical simulations of few non-Brownian droplets in various flow conditions,
specifically, focusing on their self-assembly within a short distance in a three-dimensional ($3D$) microfluidic channel, cf.\ \cite{tabeling}.
%[Shen \textit{et al., Adv. Sci.}, 2016, \textbf{3}(6):1600012].
Contrary to quasi two-dimensional ($q2D$) systems, where dipolar interaction is the key mechanism for droplet rearrangement,
droplets in $3D$ confinement produce much less disturbance to the underlying flow, thus experiencing weaker dipolar interactions.
Using confined simple shear and Poiseuille flows as reference flows,
we show that the droplet dynamics is mostly affected by the shear-induced cross-stream migration,
which favors chain structures if the droplets are under an attractive depletion force.
For more compact clusters, such as three droplets in a triangular shape,
our results suggest that an inhomogeneous cross-sectional inflow profile is further required.
Overall, the accelerated self-assembly of a small-size droplet cluster results from the combined effects of
strong depletion forces, confinement-mediated shear alignments, and fine-tuned inflow conditions.
The deterministic nature of the flow-assisted self-assembly implies the possibility of large throughputs,
though calibration of all different effects to directly produce large droplet crystals is generally difficult.
\end{paperabstract}

%------------------------------------------------------------------------------
% Article
%------------------------------------------------------------------------------
%
\input{paper2/article.tex}

%------------------------------------------------------------------------------
% Bibliography
%------------------------------------------------------------------------------
%
\clearpage
\bibliographystyle{jfm}
\bibliography{thesis}
%
\IfFileExists{paper2/paper.bbl}{\input{paper2/paper.bbl}}{}

%===============================================================================
%                            END PAPER
%===============================================================================
\end{paper}