### 2020-01-26 paper6/paper.tex

%------------------------------------------------------------------------------
% Define title, author(s), affiliation and publishing status
%
\papertitle[Effective slip over partially filled microcavities] % Short title used in headlines (optional)
{%
  Effective slip over partially filled microcavities and its possible failure% THE COMMENT SYMBOL AT THE END OF THIS LINE IS NEEDED
}%
%
\papertoctitle{Effective slip over partially filled microcavities and its possible failure} % Title for toc
%
\paperauthor[Ge, Holmgren, Kronbichler, Brandt, Kreiss] % Short authors used in headlines and List Of Papers
{%
  Zhouyang Ge$^1$, Hanna Holmgren$^2$, Martin Kronbichler$^3$, \\Luca Brandt$^1,$, Gunilla Kreiss$^2$%
}%
%
\listpaperauthor{Z. Ge, H. Holmgren, M. Kronbichler, L. Brandt, G. Kreiss}% (optional) Short authors used in List Of Papers
%
\paperaffiliation
{%
  $^1$ Linn\'e FLOW Centre and SeRC, KTH Mechanics, S-100 44 Stockholm, Sweden\\%
  $^2$ Division of Scientific Computing, Department of Information Technology, \\Uppsala University, Box 337, 751 05 Uppsala, Sweden\\%
  $^3$ Institute for Computational Mechanics, Technical University of Munich, Boltzmannstrasse 15, 85748 Garching b. München, Germany%
}%
%
\paperjournal[Phys. Rev. Fluids.] % Short publish info used in List Of Papers
{%
	Physical Review Fluids%
}%
%
\papervolume{3}%
%
\papernumber{054201}%
%
\paperpages{}%
%
\paperyear{2018}%
%
\papersummary%
{% Insert summary of the paper here (used in introduction)
   In this paper, we combine two separate simulation methods (phase field and level set methods) to numerically study
   the transverse flow over a 2D model liquid-infused surface (LIS).
   Computationally, we introduce and verifie a multiscale numerical framework for wetting modelling
   by locally solving the Cahn-Hilliard equations near the moving contact line within a larger Stokes system.
   Physically, we conduct a parametric study of the performance of the LIS, quantified by its effective slip length and potential drainage.
   Among the many parameters (the capillary number, viscosity ratio, contact angle, etc.) explored,
   we identify that filling fraction is the most sensitive parameter for the effective slip;
   the effects of other parameters are generally intertwined but weakened if the surface groove is not fully filled.
   Furthermore, we report a shear-driven failure mode associated with lubricant drainage,
   that is worse at higher lubricant viscosities if the lubricant is less viscous than the outer fluid,
   but disappears if the lubricant is more viscous.
   Overall, our study may serve as a conceptual guide for LIS design complementing experimental tests.
}%
%
\graphicspath{{paper6/}}%
%
%
%===============================================================================
%                            BEGIN PAPER
%===============================================================================
%
\begin{paper}

\makepapertitle

%------------------------------------------------------------------------------
% Abstract
%------------------------------------------------------------------------------
%
\begin{paperabstract}
\Ge{Motivated by the emerging applications of liquid-infused surfaces (LIS), we study the drag reduction and robustness of transverse flows over two-dimensional microcavities partially filled with an oily lubricant.
Using separate simulations at different scales, characteristic contact line velocities at the fluid-solid intersection are first extracted from nano-scale phase field simulations and then applied to micron-scale two-phase flows, thus introducing a multiscale numerical framework to model the interface displacement and deformation within the cavities. 
As we explore the various effects of the lubricant-to-outer-fluid viscosity ratio $\tilde{\mu}_2/\tilde{\mu}_1$, the capillary number Ca, the static contact angle $\theta_s$, and the filling fraction of the cavity $\delta$, we find that the effective slip is most sensitive to the parameter  $\delta$. The effects of $\tilde{\mu}_2/\tilde{\mu}_1$ and $\theta_s$ are generally intertwined, but weakened if $\delta < 1$.
Moreover, for an initial filling fraction $\delta =0.94$, our results show that the effective slip is nearly independent of the capillary number, when it is small. Further increasing Ca to about $0.01 \tilde{\mu}_1/\tilde{\mu}_2$, we identify a possible failure mode, associated with lubricants draining from the LIS, for $\tilde{\mu}_2/\tilde{\mu}_1 \lesssim 0.1$. Very viscous lubricants (\eg $\tilde{\mu}_2/\tilde{\mu}_1 >1$), on the other hand, are immune to such failure due to their generally larger contact line velocity.}
\end{paperabstract}

%------------------------------------------------------------------------------
% Article
%------------------------------------------------------------------------------
%
\input{paper6/article.tex}

%------------------------------------------------------------------------------
% Bibliography
%------------------------------------------------------------------------------
%
%\clearpage
\bibliographystyle{jfm}
\bibliography{thesis}
%
\IfFileExists{paper6/paper.bbl}{\input{paper6/paper.bbl}}{}

%===============================================================================
%                            END PAPER
%===============================================================================
\end{paper}