### 2020-01-27 frontmatter.tex
added /safe for recovery.

%-------------------------------------------------------------------------------
% Cover page
%-------------------------------------------------------------------------------
%
\makecoverpage

%-------------------------------------------------------------------------------
% Dedication (comment if none)
%-------------------------------------------------------------------------------
%
\dedication{
%{\it It is true that Nature begins by reasoning and ends by experience. Nevertheless, \\we must begin with experiments and try through it to discover the reason.}\\[20pt]
%Leonardo da Vinci 
{If the scientist had an infinity of time at their disposal,\\it would be sufficient to say to them,\\``Look, and look \it{carefully}.''}\\[20pt]
Henri Poincar\'e
}

%-------------------------------------------------------------------------------
% Abstract (English)
%-------------------------------------------------------------------------------
%
% COMMENT: the abstract should be limited to 2000 characters and the keywords
% to 250 characters because otherwise they will not fit in the mailing sheet.
%
\begin{abstract}
Micron to millimetre sized droplets, precisely generated or sustained in controlled environment, have great potential in myriads of engineering applications
functioning as the basic element to assemble metamaterials, deliver drugs, host surfactant, reduce friction and damp turbulence.
The interaction of droplets from pairwise to collective levels is the most important factor in controlling these processes,
yet little is known about the detailed mechanisms in various nonideal conditions.
The present thesis combines a number of studies aiming to elucidate the physical principles of droplet interactions and suspension flow
using both high- and low-fidelity numerical simulations.

We first study flow-assisted droplet assembly in microfluidic channels, seeking to harness the droplet interactions to produce photonic bandgap materials.
A novel interface-correction level set/ghost fluid method (ICLS/GFM) is developed to directly simulate liquid droplets under depletion forces.
Comparing to previous methods, ICLS/GFM conserves the global mass of each fluid using a simple mass-correction scheme,
accurately computes the surface tension and depletion forces under the same framework,
and has subsequently been applied to investigate the droplet clustering observed in a microfluidic experiment.
Our simulations, supported by theoretical derivations, suggest that the observed fast self-assembly arises from a combination of 
strong depletion forces, confinement-mediated shear alignments of the droplets, and fine-tuned inflow conditions of the microchannel.
However, the interplay of these 3D effects negates a simple droplet interaction model of parametric dependence,
rendering the design of microfluidic chips for photonic crystal fabrications elusive in practice.

The next focus of the thesis is the implementation of a minimal discrete-element lubrication/contact dynamics model for simulation of dense particle suspensions.
Here, the main ingredients include a frame-invariant, short-range lubrication model for spherical particles,
and a soft-core, stick/slide frictional contact model activated when particles overlap.
Furthermore, since contact interactions dominate at high particle concentrations,
we expect the methodology to be applicable for probing the jamming of non-spherical particles and the rheology of foams as well.

Finally, we include two miscellaneous studies concerning
the slippage property of liquid-infused surfaces and droplets statistics in a homogeneous turbulent shear flow.
Overall, results of these simulations provide detailed flow visualisations and qualitative dependence of the target functional on various governing parameters,
facilitating experimental and theoretical investigations to design more robust drag-reducing surfaces and predict droplet distributions in emulsions.

%
\keywords{droplets, suspension, multiphase flow, microfluidics, soft matter, rheology, depletion force, level set, ghost fluid, discrete element.}
%
\end{abstract}

%-------------------------------------------------------------------------------
% Abstrakt (Swedish) 
%-------------------------------------------------------------------------------
%
% COMMENT: the abstract should be limited to 2000 characters and the keywords
% to 250 characters because otherwise they will not fit in the mailing sheet.
%
\begin{abstrakt}
Droppar av mikron till millimeter, exakt genererade eller bibehållna i kontrollerad miljö, har stor potential i många olika tekniska applikationer
fungerar som det grundläggande elementet för att montera metamaterial, leverera läkemedel, värda ytaktivt medel, minska friktion och fuktig turbulens.
Interaktion mellan droppar från parvis till kollektivnivå är den viktigaste faktorn för att kontrollera dessa processer,
ändå är lite känt om de detaljerade mekanismerna i olika nonideala förhållanden.
Den här avhandlingen kombinerar ett antal studier som syftar till att belysa de fysiska principerna för droppinteraktioner och suspensionflöde
med både numeriska simuleringar med hög och låg trovärdighet.

Vi studerar först flödesassisterade droppmontering i mikrofluidkanaler och försöker utnyttja droppinteraktioner för att producera fotoniska bandgapmaterial.
En ny gränssnittskorrigeringsnivåuppsättning/spökvätskemetod (ICLS/GFM) är utvecklad för att direkt simulera vätskedroppar under utarmningskrafter.
Jämfört med tidigare metoder sparar ICLS/GFM den globala massan för varje fluid med hjälp av ett enkelt masskorrigeringsschema,
beräknar exakt ytspänningen och utarmningskrafterna under samma ram,
och har sedan applicerats för att undersöka droppklusteringen som observerats i ett mikrofluidiskt experiment.
Våra simuleringar, med stöd av teoretiska härledningar, antyder att den observerade snabba självmonteringen uppstår från en kombination av
starka utarmningskrafter, inneslutningsförmedlade skjuvlinjeringar av dropparna och fininställda inflödesförhållanden för mikrokanalen.
Men samspelet mellan dessa 3D-effekter avskaffar en enkel droppinteraktionsmodell av parametriskt beroende,
vilket gör utformningen av mikrofluidiska chips för fotoniska kristallfabrikationer som är svåra i praktiken.

Nästa fokus för avhandlingen är implementeringen av en minimal diskret-element smörjning/kontaktdynamik-modell för simulering av täta partikelsuspensioner.
Här innehåller huvudingredienserna en ram-invariant smörjningsmodell med kort räckvidd för sfäriska partiklar,
och en friktionskontaktmodell med mjukkärna, stick/slide aktiverad när partiklar överlappar varandra.
Eftersom kontaktinteraktioner dominerar vid höga partikelkoncentrationer,
vi förväntar oss att metodologin också är tillämplig för att undersöka fastklämning av icke-sfäriska partiklar och reologi för skum.

Slutligen inkluderar vi två diverse studier rörande
glidningsegenskapen hos vätska-infunderade ytor och droppstatistik i ett homogent turbulent skjuvflöde.
Sammantaget ger resultaten av dessa simuleringar detaljerade flödesvisualiseringar och kvalitativt beroende av målfunktionen på olika reglerande parametrar,
underlätta experimentella och teoretiska undersökningar för att utforma mer robusta dragreducerande ytor och förutsäga droppfördelningar i emulsioner.
%
\nyckelord{droppar, suspension, flerfasflöde, mikrofluidik, mjukt material, reologi, utarmningskraft, nivåuppsättning, spökvätska, diskret element.}
%
\end{abstrakt}

%-------------------------------------------------------------------------------
% Preface page
%-------------------------------------------------------------------------------
%
\begin{preface}
	This thesis summarizes a selection of studies on droplet interactions and suspension flow
        that may find applications in microfluidics, soft matter, and rheology.
        A brief introduction of the basic physical concepts and numerical methods is presented in Part I,
        followed by seven journal articles or preprints in Part II, see below.
        For consistency, all papers are formatted in the \texttt{jfm} style as the rest of the thesis.
        Their contents remain faithful to the original publications.
\end{preface}

%-------------------------------------------------------------------------------
% Division of work between authors
%-------------------------------------------------------------------------------
%
\begin{divisionofwork}
	The main thesis advisor is Professor Luca Brandt (LB).
	Associate Professor Outi Tammisola (OT) acts as the co-advisor.

        \bigskip

	\paperitem
                LB directed the research.
                Z. Ge (ZG) and J-Ch. Loiseau developed, implemented and validated the methods.
                OT oversaw the entire project.
		ZG wrote the paper with inputs from all other authors.

	\paperitem
                Following Paper 1, ZG continued the research, performed the simulation
                and wrote the paper with inputs from the rest of the authors.

	\paperitem
                A. Leshansky (AL) directed the research.
                I. Fouxon (IF) derived the theory.
		ZG performed the simulation.
                IF wrote the paper with inputs from the rest of the authors.

	\paperitem
                AL directed the research.
                IF constructed the theory.
		B. Rubinstein performed the multipole expansion simulation.
                ZG performed the Navier-Stokes simulation.
                IF wrote the paper with inputs from all other authors.

	\paperitem
                LB conceived the project.
                M.E. Rosti (MER) implemented the method and performed the simulation.
		MER and ZG performed the droplet statistics and scaling analyses.
                S.S. Jain performed the droplet rendering.
                M.S. Dodd performed the turbulent kinetic energy analysis.
                MER wrote the paper with inputs from the rest of the authors.

	\paperitem
                G. Kreiss initiated the project.
                H. Holmgren implemented the code and performed the preliminary Stokes simulation.
                M. Kronbichler performed the phase field simulation.
		ZG performed the final Stokes simulation and 
                wrote the paper with inputs from the rest of the authors.

	\paperitem
                LB led the research.
		ZG implemented the method and wrote the report with input from LB.

\end{divisionofwork}

%-------------------------------------------------------------------------------
% Additional publications (comment if none)
%-------------------------------------------------------------------------------
%
%\begin{otherpublications}
%	The following papers, although related, are not included in this thesis.
%
%  \paperitem%
%    {Anakin Skywalker \& Master Obi-Wan Kenobi}% Authors
%    {3639}% Year
%    {The light sabre: an elegant weapon for a more civilized age}% Title
%    {Jedi Journal of Weapons}% Journal
%    {33}% Volume
%    {2}% Number
%    {pp. 55--60}% Pages
%
%\end{otherpublications}

%-------------------------------------------------------------------------------
% Conferences (comment if none)
%-------------------------------------------------------------------------------
%
\clearpage
\begin{conferences}
        In addition to journal publications, part of the work in this thesis were
        presented in the following conferences and workshops. The presenting author is underlined.

        \bigskip

  \conferenceitem%
    {\underline{Z. Ge}, L. Brandt}% Authors
    {June 2016}% Year
    {Simulation of the self-assembly of colloidal particles in a microchannel}% Title
    {7$^{th}$ Summer school of Complex Motions in Fluids}% Conference
    {Zenderen, Twente, Netherlands}% Location

  \conferenceitem%        
    {Z. Ge, \underline{O. Tammisola}, J.Ch. Loiseau, L. Brandt}% Authors
    {September 2016}% Year
    {Direct numerical simulation of the self-assembly of colloidal particles in a micro-channel}% Title
    {1$^{st}$ International Conference on Multiscale Applications of Surface Tension}% Conference
    {Brussel, Belgium}% Location

  \conferenceitem%
    {\underline{Z. Ge}, L. Brandt}% Authors
    {November 2016}% Year
    {Do self-assembly colloidal droplets behave like droplets?}% Title
    {69$^{th}$ Annual Meeting of the APS Division of Fluid Dynamics}% Conference
    {Portland, Oregon, USA}% Location

  \conferenceitem%
    {\underline{Z. Ge}, O. Tammisola, L. Brandt}% Authors
    {June 2017}% Year
    {A fast mass-preserving interface-correction level set/ghost fluid method for colloidal suspensions under depletion forces}% Title
    {3$^{rd}$ International Conference on Numerical Methods of Multiphase Flows (ICNMMF-III)}% Conference
    {Tokyo, Japan}% Location

  \conferenceitem%
    {\underline{Z. Ge}, H. Holmgren, M. Kronbichler, L. Brandt, G. Kreiss}% Authors
    {September 2018}% Year
    {Effective slip over partially filled microcavities and its possible failure}% Title
    {12$^{th}$ European Fluid Mechanics Conference}% Conference
    {Vienna, Austria}% Location

  \conferenceitem%
    {\underline{Z. Ge}, O. Tammisola, L. Brandt}% Authors
    {November 2018}% Year
    {Flow-assisted droplet assembly in a 3D microfluidic channel}% Title
    {71$^{st}$ Annual Meeting of the APS Division of Fluid Dynamics}% Conference
    {Atlanta, Georgia, USA}% Location

  \conferenceitem%
    {\underline{Z. Ge}, O. Tammisola, L. Brandt}% Authors
    {December 2018}% Year
    {Flow-assisted droplet assembly in a 3D microfluidic channel}% Title
    {Scattering and Dynamics of Flowing Soft Material}% Conference
    {Lund, Sweden}% Location

  \conferenceitem%
    {\underline{Z. Ge}, O. Tammisola, L. Brandt}% Authors
    {February 2019}% Year
    {Flow-assisted droplet assembly in a 3D microfluidic channel}% Title
    {Colloidal Science \& Metamaterials (CSM)}% Conference
    {Paris, France}% Location

\end{conferences}

%-------------------------------------------------------------------------------
% Table of contents
%-------------------------------------------------------------------------------
%
\tableofcontents