### 2020-02-01 acknowledgements.tex
acknowledgements added

\begin{acknowledgements}
%  Advisors,
%  funders,
%  collaborators,
%  mentors,
%  colleagues,
%  friends,
%  family.

  This thesis was written from December 2019 to February 2020, during one of the warmest winters over the last five years in Stockholm.
  Looking back, I have encountered, learned and benefited from so many people that is nearly impossibly to name them one by one.
  I am grateful to \emph{everyone} and apologize in advance for any negligence.

  First and foremost, I want to thank Professor Luca Brandt for accepting me in and guiding me through the PhD program at KTH Mechanics.
  His strong leadership from the inception,
  enthusiastic responses and constructive criticisms at every discussion,
  and the increasing freedom and trust given to me towards the end are indispensible for me to develop academically.
  For these reasons, I owe my deepest gratitude to him.

  Second, I wish to extend my sincere thanks to my day-to-day advisors and mentors at KTH.
  I thank Outi Tammisolar for co-advising me since my first project, following my progress throughout the PhD,
  and offering me invaluable career advices in the later years.
  I thank Jean-Christophe Loiseau for teaching me proper scientific programming and instilling a mathematical mindset into me.
  I also thank Mehdi Niazi for always being available for technical discussions or even code debugging, and sharing uncountable personal advices.
  The patience, support and encouragemnt they offered me are instrumental in building my confidence and independence, which I will never take for granted.

  Third, I want to express my appreciation to all collaborators and senior colleagues that has shaped and contributed to my work in the last few years.
  These include, but are not limited to, international collaborators: Michael Dodd, Alexender Leshansky, Itzhak Fouxon, and Naoki Takeishi;
  former and present colleagues at KTH: Walter Fornari, Pedro Costa, Francesco de Vita, Marco Rosti, B. M. Ningegowda, Ugis Lacis;
  office mates: Ricardo Vinuesa, Ekaterina Ezhova, Tímea Kékesi, Arash Alizad Banaei, Ashwin Vishnu;
  and the professors that I enjoy talking to or playing innebandy with:
  Shervin Bagheri, Christophe Duwig, Hanno Essén, Nicholas Apazidis, Lanie Gutierrez-Farewik, Fredrik Lundell.
  Many others have also helped, corrected or inspired me in different research projects; see the separate acknowledgements after each paper for details.

  In his famous essay, \emph{The World As I See It}, Einstein wrote,
  ``But from the point of view of daily life, without going deeper, we exist for our fellowmen --
  in the first place for those on whose smiles and welfare all our happiness depends,
  and next for all those unknown to us personally with whose destinies we are bound up by the tie of sympathy.''
  In the last several years, I have been lucky to derive plentiful happiness and sympathy from Artem, Erik, Guillaume, Nicolas, Frida, Natasha, Thea, Lee,
  as well as the rest of the Rosagstullsbacken corridor.
  Particularly, I am blessed to meet Dia, whose courage, ambition, curiosity and sensitivity have become a continuing source of my drive,
  making me feel attached and complete.

  Finally, none of the above would have been possible without the love, education, protection and sacrifice from my family at home.
  \chinese{爸，妈，感谢你们对我从小到大的培养。我的一切都属于你们。}

\end{acknowledgements}