### 2020-02-26 paper7/paper.tex

%------------------------------------------------------------------------------
% Define title, author(s), affiliation and publishing status
%
\papertitle[Note of HLGD for dense suspensions] % Short title used in healines (optional)
{%
 Implementation note on a minimal hybrid lubrication/granular dynamics model\\for dense suspensions% THE COMMENT SYMBOL AT THE END OF THIS LINE IS NEEDED
}%
%
\papertoctitle{Implementation note on a minimal hybrid lubrication/\\granular dynamics model for dense suspensions} % Title for toc
%
\paperauthor[Ge \& Brandt] % Short authors used in headlines and List Of Papers
{%
  Zhouyang Ge \& Luca Brandt%
}%
%
\listpaperauthor{Z. Ge, L. Brandt}% (optional) Short authors used in List Of Papers
%
\paperaffiliation
{%
  Linn\'e FLOW Centre and SeRC, Department of Engineering Mechanics,\\
  KTH Royal Institute of Technology, S-100 44 Stockholm, Sweden%
}%
%
\paperjournal[Technical Report] % Short publish info used in List Of Papers
{%
	Technical Report%
}%
%
\papervolume{}%
%
\papernumber{}
%
\paperpages{}%
%
\paperyear{2020}%
%
\papersummary%
{% Insert summary of the paper here (used in introduction)
   This report documents the implementation detail of a minimal hybrid lubrication/granular dynamics model
   intended for simulation of dense (potentially nonspherical) particle suspensions in shear flows.
   The method is meshless, explicit and solves directly the particle dynamics in the overdamped limit
   by balancing conservative and dissipative interactions at the pair level.
   Potential sources of inaccuracy include omissions of long-range hydrodynamic interactions and finite particle inertia due to numerical integration.
   The implementation is tested for $200\sim 500$ particles under volume fractions from 40\% to 64\%,
   showing jamming transition, continuous and discontinuous shear thickening that are in excellent agreement with literature results.
}%
%
\graphicspath{{paper7/}}%
%
%
%===============================================================================
%                            BEGIN PAPER
%===============================================================================
%
\begin{paper}

\makepapertitle

%------------------------------------------------------------------------------
% Abstract
%------------------------------------------------------------------------------
%
\begin{paperabstract}
	We describe and summarize a class of minimal numerical models emerged from recent development of simulation methods for
        dense particle suspensions in overdamped linear flows.
        The main ingredients include
        (i) a frame-invariant, short-range lubrication model for spherical particles, and
        (ii) a soft-core, stick/slide frictional contact model activated when particles overlap.
        We implement a version of the model using a modified velocity-Verlet algorithm that
        explicitly solves the $N$-body dynamical system in $\mathcal{O}(cN)$ operations,
        where $c$ is a kernel constant depending on the cutoff of particle interactions.
        The implementation is validated against literature results on jamming transition and shear thickening suspensions from 40\% to 64\% volume fractions.
        Potential strategies to extend the present methodology to non-spherical particles are also suggested for very concentrated suspensions.

\end{paperabstract}

%------------------------------------------------------------------------------
% Article
%------------------------------------------------------------------------------
%
\input{paper7/article.tex}

%------------------------------------------------------------------------------
% Bibliography
%------------------------------------------------------------------------------
%
%\clearpage
\bibliographystyle{jfm}
\bibliography{thesis}
%
\IfFileExists{paper7/paper.bbl}{\input{paper7/paper.bbl}}{}

%===============================================================================
%                            END PAPER
%===============================================================================
\end{paper}